---
title: "Untitled"
author: "Tao"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r}
# install.packages("xlsx")
library("openxlsx")
library("readxl")
#install.packages("strucchange")
library("strucchange")
library("lubridate")
library("dplyr") # for data wrangling and pipelines
library("ggplot2") # for modifying fasstr plots
# install.packages("qqplotr")
library(qqplotr)
#The R package rnoaa is used to access NOAA NCDC climate data.
#install.packages("rnoaa")
library("rnoaa")
library("rgdal")
library("maptools")
library("utils")
# install.packages("ppcc")
library("ppcc")
# install.packages("gbm")
library("gbm")
# install.packages("caret")
library("caret")
#install.packages("pdp")
library("pdp")          # model visualization
```

```{r}
#unzip("goat_creek.zip")
getwd()
ws<- readOGR("layers/globalwatershed.shp")
extent<-raster::extent(ws)
```

```{r}
#This function returns an object with a dataframe with meta-information about all available GHCND weather stations.
#stations<-ghcnd_stations()
load("station.Rdata")
stations
#save(stations , file = "station.Rdata")
ws_stations<-stations[stations$latitude> extent@ymin &
           stations$latitude<extent@ymax&
           stations$longitude>extent@xmin&
           stations$longitude<extent@xmax &
             stations$last_year>2019,]
unique(ws_stations$id)
```

#PRISM_ppt_tmin_tmean_tmax_tdmean_vpdmin_vpdmax_early_4km_20191001_20230401_44.2192_-114.9517-GC
```{r}
prism<-read.csv("PRISM_ppt_tmin_tmean_tmax_tdmean_vpdmin_vpdmax_early_4km_20191001_20230401_44.2192_-114.9517-GC.csv", skip = 10)
prism$Date <-as.Date(prism$Date, format="%Y-%m-%d")
summary(prism)
head(prism)
```

```{r}
xlsx="FY21_CBWTP_Monitoring_IWRB_Temp_data_01.13.2022.xlsx"
sheets<-excel_sheets("FY21_CBWTP_Monitoring_IWRB_Temp_data_01.13.2022.xlsx")
sheets
```

```{r}
gc<-read_excel(xlsx , sheet = sheets[9], skip = 11 )
gc$water_t_c<-   (gc$`Temperature (f)` -32)*5/9
gc$Date_1<-as.Date(gc$Date)
gc<-merge(gc, prism, by.x= "Date_1",by.y= "Date")
summary(gc)
head(gc)
plot(gc$Date, gc$vpdmax..hPa.)
```

```{r}
#saveRDS(gc,file="gc.rds")
```

```{r}
plot(gc$Date, gc$ppt..mm.   )
ggplot( gc, aes(  Date, ppt..mm.  )) +
    geom_bar(stat="identity", na.rm = TRUE) +
    ggtitle("Daily Precipitation") +
    xlab("Date") + ylab("Precipitation (mm)") +#    scale_x_date(labels=date_format ("%b %y"), breaks=date_breaks("1 year")) +
    theme(plot.title = element_text(lineheight=.8, face="bold", size = 20)) +   theme(text = element_text(size=18))
```

```{r}
hist(gc$`Temperature (f)`)
hist(gc$tmean..degrees.C.)
hist(gc$water_t_c )
hist(gc$`Flow (cfs)`)
hist(log(gc$`Flow (cfs)`))
```

The Q-Q plot, or quantile-quantile plot, is a graphical tool to help us assess if a set of data plausibly came from some theoretical distribution such as a Normal or exponential.

```{r}
ppccTest( gc$`Flow (cfs)`, "qnorm") #Normal
## see Table 2 of Looney and Gulledge (1985, p.78)
ggplot(mapping = aes(sample =  gc$`Flow (cfs)` )) + stat_qq_point(size = 2)
ggplot(mapping = aes(sample =  gc$`Flow (cfs)` ))+ stat_qq_point(size = 2,color = "red")+ stat_qq_line(color="dodgerblue")+ xlab("x-axis") + ylab("y-axis")
```

Construct F-statistics to consider whether or not a change occurred at time use the residuals for a subsample are compared to the residuals for the unsegmented dataset, 

#breakpoint of air_t
```{r}
myts <- ts(  gc$tmean..degrees.C. , start = c( 1 #julian( morgan$Date[1]) 
                                             ), frequency = 1)
myts
plot(myts)
fs.myts <- Fstats(myts ~ 1)
plot(fs.myts)
breakpoints(fs.myts)
lines(breakpoints(fs.myts))
breakpoints<-breakpoints(fs.myts)$breakpoints
plot(gc$Date, gc$tmean..degrees.C.)
par(new=T)
abline(v=gc[breakpoints,])
#plot(morgan[breakpoints,]$Date, morgan[breakpoints,]$tmean..degrees.C., col="red")
plot(gc$Date, (gc$tmean..degrees.C. ) , type="l", col="dodgerblue", xlab="", ylab="Air temperature (degC) ", main="Goat Creek")
abline(v=gc[breakpoints,], col="red")
```

#breakpoint of  
```{r}
myts <- ts(  gc$`Flow (cfs)`, start = c( 1 #julian( morgan$Date[1]) 
                                             ), frequency = 1)
myts
plot(myts)
fs.myts <- Fstats(myts ~ 1)
plot(fs.myts)
breakpoints(fs.myts)
lines(breakpoints(fs.myts))
breakpoints<-breakpoints(fs.myts)$breakpoints
plot(gc$Date, log(gc$`Flow (cfs)`) , type="l", col="dodgerblue", xlab="", ylab="log(Flow (cfs))", main="Goat Creek")
abline(v=gc[breakpoints,], col="red")
```

```{r}
myts <- ts(  gc$`Temperature (f)` , start = c( 1 #julian( morgan$Date[1]) 
                                             ), frequency = 1)
myts
plot(myts)
fs.myts <- Fstats(myts ~ 1)
plot(fs.myts)
breakpoints(fs.myts)
lines(breakpoints(fs.myts))
breakpoints<-breakpoints(fs.myts)$breakpoints
#plot(bsc$Date,  bsc$`Temperature (f)`   )
#abline(v=bsc[breakpoints,])
plot(gc$Date,  gc$water_t_c , type="l", col="dodgerblue", xlab="", ylab="Stream temperature (degC)", main="Goat Creek")
abline(v=gc[breakpoints,], col="red")

```

```{r}
summary(lm( gc$water_t_c~gc$`Flow (cfs)`+ gc$tmean..degrees.C.))
x=seq(-5,30)
plot(gc$tmean..degrees.C., gc$water_t_c, xlim=c(-5,25), ylim=c(-5,25) )
par(new=T)
plot(x,x, xlim=c(-5,25), ylim=c(-5,25),type="l")
```

```{r}
gc_1<-gc[!is.na(gc$`Flow (cfs)`) , c("water_t_c","Flow (cfs)", "tmean..degrees.C.")]
# set seed for generating random data.
set.seed(1)          
# createDataPartition() function from the caret package to split the original dataset into a training and testing set and split data into training (80%) and testing set (20%)
parts = createDataPartition(  gc_1$water_t_c , p = .75, list = F)
train =gc_1[parts,c( "water_t_c","Flow (cfs)", "tmean..degrees.C." ) ]
test = gc_1[-parts,c( "water_t_c", "Flow (cfs)", "tmean..degrees.C.")  ]
# feature and target array
test_x = test[, -1] 
test_y = test[, 1] 
```

Fits generalized boosted regression models.
```{r}
model_gbm = gbm( train$water_t_c     ~.,
                data = train,
                distribution = "gaussian",
                cv.folds = 10,
                shrinkage = .01,
                n.minobsinnode = 10,
                n.trees = 800)
 
# model performance
perf_gbm1 = gbm.perf( model_gbm, method = "cv")
print(model_gbm)
summary(model_gbm)
rinf<-summary(model_gbm)
barplot( rinf$rel.inf , horiz = TRUE, las = 1)
ggplot(rinf, aes(rel.inf)) + geom_bar()
ggplot( rinf, aes( var , rel.inf ))+  geom_col()+ 
  coord_flip()
```

```{r}
#rinf$var<- factor(rinf$var, levels=c(  "max_AirTemperature_C" ,"log_median_Q"
             #,"min_Discharge"             ,      "max_SolarRadiation" ,
#             "doy"  ))
ggplot( rinf, aes( var , rel.inf ))+  geom_col()+ 
  coord_flip()
ggplot( rinf )+  geom_bar(  aes( x=var, y= rel.inf),  stat = "summary")+ scale_x_discrete(labels=  c( "Streamflow", "mean Air Temperature C" ) )+ylab("Relative importance (%)") +xlab(" ")  + theme(text=element_text(size=14))
```

```{r}
#test_y <-test_y$max_StreamTemp
pred_y = predict.gbm(model_gbm, test_x)
residuals =  test_y   - pred_y
summary(test_y )
xlim=c(5,30)
RMSE = sqrt(mean(residuals^2))
cat('The root mean square error of the test data is ', round(RMSE,3),'\n')
y_test_mean = mean( test_y  )
# Calculate total sum of squares
tss =  sum(( test_y   - y_test_mean)^2 )
# Calculate residual sum of squares
rss =  sum(residuals^2)
# Calculate R-squared
rsq  =  1 - (rss/tss)
cat('The R-square of the test data is ', round(rsq,3), '\n')
# visualize the model, actual and predicted data
x_ax = 1:length(pred_y)
plot(x_ax, test_y  , col="blue", pch=20, cex=.9)
lines(x_ax, pred_y, col="red", pch=20, cex=.9) 
plot(  test_y , pred_y,xlim= xlim ,ylim= xlim, xlab="Observed stream temperature (degC)", ylab="Predicted stream temperature (degC)", main="GBM")
par(new=T)
x=seq(1,30)
plot(x,x,type="l",xlim= xlim ,ylim= xlim,xlab="",ylab="")
```
